Supplementary Materials- A Bayesian approach to burial networks to infer social changes in northeastern Taiwan during the European colonization period

### 2021-01-26 analysis/paper/supplementary-material.Rmd
merge conflicts and edit SI

-   further historical background of northeastern Taiwan in a pericolonial context
-   further definition of our approach to tie formation with the beads from Kiwulan burials
-   further review of network analysis in archaeology
-   further details on methods of modeling fitting, and
-   assessment of MCMC output and Goodness-of-fit diagnostics for our models

# Kiwulan in the 17th northern Taiwan {-}

@@

## Tie formation using bead data from Kiwulan

@@

Current approaches to network analysis used by archaeologists are mostly restricted to a single rational structure without consideration of interaction between network variables. Our use of Bayesian inference on ERGMs is the first application to archaeological data that can bring new insights to understand past social structures by characterizing network properties as a whole. ERGMs are an important family of statistical models for networks that allows direct modeling for the formation of edges, or ties, between nodes [@Robins2007].

The assumption is that possible ties in a network are random variables and dependent on actor variables or the presence or absence of other ties [@Robins2007]. In other words, networks in ERGMs are viewed as dependent variables, where network dependencies and the attributes of nodes/edges can influence the formation of a tie [@Snijders2011]. For example, nodes with similar attributes are more likely to form a relationship, such as friendship between people with the same hobby. Ties form a small structure in a network called a graph configuration, that describes the form of dependence, such as reciprocity (relationship between two actors), transitivity or clustering (relationship between two actors through a shared third actor), homophily (relationship between actors with a similar attribute), and popularity (actors have many relationships with others) [@Robins2007; @Snijders2006; @Morris2008]. Those configurations represent the structure or the property of a network and can be expressed by network statistics.

@@

After we set our model parameters, we simulated networks in a Bayesian framework using a Markov chain Monte Carlo (MCMC) algorithm. MCMC algorithms allow estimation of posterior distributions through direct random sampling the posterior without assuming the prior comes from any specific distribution [@Hamra2013]. We can obtain a posterior distribution by constructing a Markov chain that describes a sequence of moves from current state to the next state following probabilistic rules based on an algorithm. This enables a random or stochastic simulation in a long run where each move does not depend on the previous move. More chains ensures a more desirable posterior distribution that is close to the target distribution under study, or convergence.

In Bayesian ERGMs, MCMC first selects a set of edges (or a set of empty pair of actors) with equal probability, and then switches to a pair of actors at random within the chosen set [@Caimo2011]. In our case, we set the number of chains to six. For each chain, the number of burn-in iterations was `r burn_in` and the number of iterations after the burn-in was `r main_iter`. We set the number of iterations used to simulate a network y' at each iteration to `r aux_iter`.

@@

## Assessment of MCMC output and Goodness-of-fit diagnostics for our models

@@

### Colophon

>>>>

-  further historical background of 17th century northeastern Taiwan in a pericolonial context
-  further definition of our approach to tie formation with the beads from Kiwulan burials
-  further review of network analysis and archaeological applications
-  further details on methods of modeling fitting, and
-  assessment of MCMC output and Goodness-of-fit diagnostics for our models

## Kiwulan in the 17th century northeastern Taiwan {.unnumbered}

@@

## Tie formation with the beads from Kiwulan burials {.unnumbered}

@@

Current approaches to network analysis used by archaeologists are mostly restricted to a single rational structure without consideration of interaction between network variables. Our use of Bayesian inference on ERGMs is the first application to archaeological data that can bring new insights to understand past social structures by characterizing network properties as a whole. ERGMs are an important family of statistical models for networks that allows direct modeling for the formation of edges, or ties, between nodes [@Robins2007]. The assumption is that possible ties in a network are random variables and dependent on actor variables or the presence or absence of other ties [@Robins2007]. In other words, networks in ERGMs are viewed as dependent variables, where network dependencies and the attributes of nodes/edges can influence the formation of a tie [@Snijders2011]. For example, nodes with similar attributes are more likely to form a relationship, such as friendship between people with the same hobby. Ties form a small structure in a network called a graph configuration, that describes the form of dependence, such as reciprocity (relationship between two actors), transitivity or clustering (relationship between two actors through a shared third actor), homophily (relationship between actors with a similar attribute), and popularity (actors have many relationships with others) [@Robins2007; @Snijders2006; @Morris2008]. Those configurations represent the structure or the property of a network and can be expressed by network statistics.

@@

After we set our model parameters, we simulated networks in a Bayesian framework using a Markov chain Monte Carlo (MCMC) algorithm. MCMC algorithms allow estimation of posterior distributions through direct random sampling the posterior without assuming the prior comes from any specific distribution [@Hamra2013]. We can obtain a posterior distribution by constructing a Markov chain that describes a sequence of moves from current state to the next state following probabilistic rules based on an algorithm. This enables a random or stochastic simulation in a long run where each move does not depend on the previous move. More chains ensures a more desirable posterior distribution that is close to the target distribution under study, or convergence. In Bayesian ERGMs, MCMC first selects a set of edges (or a set of empty pair of actors) with equal probability, and then switches to a pair of actors at random within the chosen set [@Caimo2011]. In our case, we set the number of chains to six. For each chain, the number of burn-in iterations was `r burn_in` and the number of iterations after the burn-in was `r main_iter`. We set the number of iterations used to simulate a network y' at each iteration to `r aux_iter`.

@@

## Assessment of MCMC output and Goodness-of-fit diagnostics for our models {.unnumbered}

@@

### Colophon {.unnumbered}
